### 2021-04-25 submission/manuscript.tex
Add references

Osmotic laxatives can lead to diarrhea depending on the administered
dose and temporarily disrupt the human intestinal microbiota (13). The
ubiquitous osmotic laxative, polyethylene glycol (PEG) 3350 is found in
Miralax, Nulytely, and Golytely and is also commonly used as bowel
preparation for colonoscopies. Interestingly, previous studies have
shown that treating mice with PEG alone altered microbiota composition,
reduced acetate and butyrate production, altered the mucus barrier, and
rendered the mice susceptible to \emph{C. difficile} infection (14--17).
The mucus barrier is thought to mediate protection from \emph{C.
difficile} infections by protecting intestinal epithelial cells from the
toxins produced by \emph{C. difficile} (Ref). However, whether laxative
results in more severe CDIs in mice and how long mice remain colonized
with \emph{C. difficile} after challenge is unclear.

Beyond susceptibility, PEG is also relevant in the context of treating
recurrent CDIs via fecal microbiota transplant (FMT) where a healthy
microbiota is administered to the patient to restore colonization. For
FMTs that are delivered via colonoscopy, patients typically undergo
bowel preparation by taking an osmotic laxative prior to the procedure.
Many of the FMT studies to date rationalize the use of laxatives (Ref)
based on a 1996 case study with 2 pediatric patients where the authors
suggested in the discussion that the laxative may help flush \emph{C.
difficile} spores and toxins from the intestine (18).

In the past, our group has used C57BL6 mice to characterize how
antibiotics including clindamycin disrupt the microbiota and influence
\emph{C. difficile} susceptibility and clearance {[}ref{]}. Although,
two groups have now shown PEG treatment alone renders mice susceptible
to \emph{C. difficile}, these studies have raised additional questions
regarding the dynamics and severity of infection as well as the role of
laxative treatment in \emph{C. difficile} clearance that should be
addressed to better inform how we think about laxatives in the context
of CDIs. Here, we used our C57BL/6 clindamycin model as a control group
to characterize how long PEG-treated mice remain susceptible, whether
PEG treatment results in more severe CDI and sustained \emph{C.
difficile} colonization, and if PEG treatment post-CDI can promote
\emph{C. difficile} clearance.

@@

\textbf{5-day laxative treatment leads to prolonged \emph{C. difficile}
colonization in mice.} We compared mice treated with the osmotic
laxative PEG 3350 to our standard 10 mg/kg clindamycin treatment, which
temporarily renders the mice susceptible to \emph{C. difficile}, with
mice typically clearing \emph{C. difficile} within 10 days
post-infection (9, 19). All PEG-treated mice were administered a 15\%
PEG solution in the drinking water for 5-days, one group was also
treated with clindamycin, and one group was allowed to recover for 10
days prior to challenge (Fig. 1A). PEG treatment resulted in weight loss
in all 3 groups of PEG-treated mice, with the greatest change in weight
observed on the fifth day of PEG treatment and the mice regained most of
the weight five days after treatment (Fig. 1B). After either PEG,
clindamycin, or PEG and clindamycin treatment all mice were challenged
with 10\textsuperscript{3} \emph{C. difficile} 630 spores (Fig. 1A). All
treatments rendered mice susceptible to \emph{C. difficile}
colonization. However, PEG-treated mice remained colonized with \emph{C.
difficile} at high levels through thirty days post-infection (Fig. 1C).
In contrast, the clindamycin-treated mice cleared \emph{C. difficile}
within ten days post-infection. Surprisingly, PEG-treated mice were
still susceptible to \emph{C. difficile} infection after 10-days of
recovery from treatment although \emph{C. difficile} was not detectable
in most of the group in the initial five days post-infection (Fig. 1C).
From 9 days post-infection onward, the median \emph{C. difficile}
stabilized for the 10-day recovery group of PEG-treated mice and
remained high through 30 days post-infection (Fig. 1C). Thus, osmotic
laxative treatment alone was sufficient to render mice susceptible to
prolonged \emph{C. difficile} colonization and PEG-treated mice remained
susceptible for up to ten days post-treatment.

\textbf{5-day laxative treatment differentially disrupts the fecal
microbiota compared to clindamycin treatment.} Since osmotic laxatives
and clindamycin have previously been shown to disrupt the murine
microbiota (ref), we hypothesized the different \emph{C. difficile}
colonization dynamics between mice treated with the osmotic laxative or
clindamycin were due to the two drugs having differential effects on the
microbiota. We profiled the stool microbiota over time by sequencing the
V4 region of the 16S rRNA gene to compare changes across treatment
groups. We found time and treatment group explained half of the observed
variation between fecal communities with most of the remaining variation
explained by interactions between treatment group and other experimental
variables including time, cage effects, and sequencing preparation
(PERMANOVA combined R\textsuperscript{2} = 0.95, \emph{P} \textless{}
0.001, Fig. 2A, Data Set S1, Sheet X). Cage effects refer to the
well-documented phenomenon that mice housed in the same cages have
similar microbial communities due to copraphagy (ref), we tried to
minimize the impact of cage effects on our experiment by breaking up
cagemates when assigning mice to treatment groups and primarily housing
only two mice per cage. Importantly, although we conducted a total of 5
separate experiments, the experiment number and it's interaction with
treatment group was not one of the variables that significantly
explained the observed variation in fecal communities (Data Set S1,
Sheet X). Interestingly, none of the treatment groups recovered to their
baseline community structure either 10 or 30 days post-infection
suggesting other community features besides recovery to baseline were
responsible for the prolonged \emph{C. difficile} colonization in
PEG-treated mice (Fig. 2B).

@@

Next, we examined CDI severity by evaluating cecum and colon
histopathology (20) and found there was no difference in cecum and colon
scores between clindamycin and PEG-treated mice that were challenged
with \emph{C. difficile} at 4 days post-infection (Fig. 3D), the
timepoint typically examined in \emph{C. difficile} 630 challenged mice
(ref). We also looked at 6 days post-infection because that was when we
started to see a large difference in C\emph{. difficile} colonization
levels between PEG- and clindamycin-treated mice (Fig. 1C). Although,
there was a slight difference in the colon between PEG and
clindamycin-treated mice, there was no difference in the cecum and the
overall score is still relatively low given that the max possible
summary score is 12 (Fig. 3E). Thus, although PEG treatment had a
profound impact on the mucosal microbiota, the impact of \emph{C.
difficile} 630 infection on the cecum and colon was similar between PEG
and clindamycin treated mice.

\textbf{\emph{C. difficile} challenge does not have a synergistic
disruptive effect on the microbiota of PEG-treated mice} Because
\emph{C. difficile} itself can have an impact on the microbiota (21), we
also sequenced the tissue and stools of mock-challenged clindamycin and
5-day PEG treated mice. Examining the stools of the mock-challenged mice
revealed similar bacterial disruptions as the \emph{C. difficile}
challenged mice (Fig. S2A-C). Similarly, there was no difference between
the tissues of mock and \emph{C. difficile} challenged mice (Fig.
S2D-F). Thus, most of the microbiota alterations we observed in the
PEG-treated mice were a result of the laxative and not an interaction
between the laxative and \emph{C. difficile}.

@@

\textbf{Five-day post-infection community data can predict mice that
will have prolonged \emph{C. difficile} colonization} After identifying
bacteria associated with the 5-day, 1-day and post-CDI 1-day PEG
treatments, we decided to examine the taxa that were influencing
prolonged \emph{C. difficile} colonization. We trained 3 types of
machine learning models (random forest, logistic regression, and support
vector machine) with input bacterial community data from 5 days
post-infection to predict whether the mice were colonized with \emph{C.
difficile} 10 days post-infection. We chose 5 days post-infection
because that was the earliest timepoint where we would see a treatment
effect in the mice that were given 1-day PEG treatment three days after
\emph{C. difficile} challenge and then administered an FMT or PBS
gavage. The random forest model had the highest performance (AUROC =
0.90, Data Set S1, Sheet X), so we next performed permutation importance
to examine the bacteria that were driving performance. We selected the
top 10 bacteria contributing to our models performance (Fig. 7A) and
examined their relative abundance at 5 days post-infection, the
timepoint used to predict C. difficile colonization status on day 10
(Fig. 7B). Next, we focused on the 5 genera that were \textgreater{} 1
\% relative abundance in either the cleared or colonized mice and
examined the relative abundance dynamics of these bacteria over time. We
found \emph{, }, and \_ tended to have a higher relative abundance and
\textbf{, } had a lower relative abundance in the mice with prolonged
colonization compared to the mice that cleared \emph{C. difficile} (Fig.
7C). Previous work examining the impact of PEG on the murine microbiota
found that PEG treatment resulted in the permanent loss of S24-7, also
known as \emph{Muribaculum intestinale} (16). We decided to check our
\emph{Porphyromonadaceae} OTUs because \emph{Muribaculum intestinale} is
known to be classified as \emph{Porphyromonadaceae} by the Ribosomal
Database Project (RDP) database (22), \emph{Porphyromonadaceae} was a
top feature in the random forest model predicting prolonged \emph{C.
difficile} colonization, and had a high relative abundance in the
communities of mice that cleared \emph{C. difficile} within 10 days. We
identified 4 OTUs that had at least (insert minimum percent identity) to
\emph{Muribaculum intestinale} and examined their abundance in mice that
either cleared or were still colonized with \emph{C. difficile} at 10
days post-infection (Fig. S4).

@@

  \begin{itemize}
  \tightlist
  \item
    What's known regarding laxatives and susceptibility to CDIs
  \item
    Relevance to human FMTs? Unclear what the best administration route
    is because there have been no studies designed to evaluate the best
    administration route for FMTs.
  \end{itemize}
\end{itemize}

@@

@@

\leavevmode\hypertarget{ref-Schubert2015}{}%
12. \textbf{Schubert AM}, \textbf{Sinani H}, \textbf{Schloss PD}. 2015.
Antibiotic-induced alterations of the murine gut microbiota and
subsequent effects on colonization resistance against \emph{Clostridium
difficile}. mBio \textbf{6}.
doi:\href{https://doi.org/10.1128/mbio.00974-15}{10.1128/mbio.00974-15}.

@@

\leavevmode\hypertarget{ref-Liacouras1996}{}%
18. \textbf{Liacouras CA}, \textbf{Piccoli DA}. 1996. Whole-bowel
irrigation as an adjunct to the treatment of chronic, relapsing
clostridium difficile colitis. Journal of Clinical Gastroenterology
\textbf{22}:186--189.
doi:\href{https://doi.org/10.1097/00004836-199604000-00007}{10.1097/00004836-199604000-00007}.

\leavevmode\hypertarget{ref-Tomkovich2020}{}%
19. \textbf{Tomkovich S}, \textbf{Stough JMA}, \textbf{Bishop L},
\textbf{Schloss PD}. 2020. The initial gut microbiota and response to
antibiotic perturbation influence clostridioides difficile clearance in
mice. mSphere \textbf{5}.
doi:\href{https://doi.org/10.1128/msphere.00869-20}{10.1128/msphere.00869-20}.

\leavevmode\hypertarget{ref-Reeves2011}{}%
20. \textbf{Reeves AE}, \textbf{Theriot CM}, \textbf{Bergin IL},
\textbf{Huffnagle GB}, \textbf{Schloss PD}, \textbf{Young VB}. 2011. The
interplay between microbiome dynamics and pathogen dynamics in a murine
model of \emph{Clostridium difficile} infection \textbf{2}:145--158.
doi:\href{https://doi.org/10.4161/gmic.2.3.16333}{10.4161/gmic.2.3.16333}.

\leavevmode\hypertarget{ref-Jenior2018}{}%
21. \textbf{Jenior ML}, \textbf{Leslie JL}, \textbf{Young VB},
\textbf{Schloss PD}. 2018. \emph{Clostridium difficile} alters the
structure and metabolism of distinct cecal microbiomes during initial
infection to promote sustained colonization. mSphere \textbf{3}.
doi:\href{https://doi.org/10.1128/msphere.00261-18}{10.1128/msphere.00261-18}.

\leavevmode\hypertarget{ref-Vornhagen2020}{}%
22. \textbf{Vornhagen J}, \textbf{Bassis CM}, \textbf{Ramakrishnan S},
\textbf{Hein R}, \textbf{Mason S}, \textbf{Bergman Y}, \textbf{Sunshine
N}, \textbf{Fan Y}, \textbf{Timp W}, \textbf{Schatz MC}, \textbf{Young
VB}, \textbf{Simner PJ}, \textbf{Bachman MA}. 2020. A plasmid locus
associated with klebsiella clinical infections encodes a
microbiome-dependent gut fitness factor.
doi:\href{https://doi.org/10.1101/2020.02.26.963322}{10.1101/2020.02.26.963322}.

>>>>

Osmotic laxatives can lead to diarrhea depending on the administered
dose and temporarily disrupt the human intestinal microbiota (13). The
ubiquitous osmotic laxative, polyethylene glycol (PEG) 3350 is found in
Miralax, Nulytely, and Golytely and is also commonly used as bowel
preparation for colonoscopies. Interestingly, previous studies have
shown that treating mice with PEG alone altered microbiota composition,
reduced acetate and butyrate production, altered the mucus barrier, and
rendered the mice susceptible to \emph{C. difficile} infection (14--17).
The mucus barrier is thought to mediate protection from \emph{C.
difficile} infections by protecting intestinal epithelial cells from the
toxins produced by \emph{C. difficile} (18, 19). However, whether
laxative results in more severe CDIs in mice and how long mice remain
colonized with \emph{C. difficile} after challenge is unclear.

Beyond susceptibility, PEG is also relevant in the context of treating
recurrent CDIs via fecal microbiota transplant (FMT) where a healthy
microbiota is administered to the patient to restore colonization. For
FMTs that are delivered via colonoscopy, patients typically undergo
bowel preparation by taking an osmotic laxative prior to the procedure.
Many of the FMT studies to date rationalize the use of laxatives
(20--22) based on a 1996 case study with 2 pediatric patients where the
authors suggested in the discussion that the laxative may help flush
\emph{C. difficile} spores and toxins from the intestine (23).

In the past, our group has used C57BL6 mice to characterize how
antibiotics including clindamycin disrupt the microbiota and influence
\emph{C. difficile} susceptibility and clearance (24--26). Although, two
groups have now shown PEG treatment alone renders mice susceptible to
\emph{C. difficile} (15, 17), these studies have raised additional
questions regarding the dynamics and severity of infection as well as
the role of laxative treatment in \emph{C. difficile} clearance that
should be addressed to better inform how we think about laxatives in the
context of CDIs. Here, we used our C57BL/6 clindamycin model as a
control group to characterize how long PEG-treated mice remain
susceptible, whether PEG treatment results in more severe CDI and
sustained \emph{C. difficile} colonization, and if PEG treatment
post-CDI can promote \emph{C. difficile} clearance.

@@

\textbf{5-day laxative treatment leads to prolonged \emph{C. difficile}
colonization in mice.} We compared mice treated with the osmotic
laxative PEG 3350 to our standard 10 mg/kg clindamycin treatment, which
temporarily renders the mice susceptible to \emph{C. difficile}, with
mice typically clearing \emph{C. difficile} within 10 days
post-infection (9, 26). All PEG-treated mice were administered a 15\%
PEG solution in the drinking water for 5-days, one group was also
treated with clindamycin, and one group was allowed to recover for 10
days prior to challenge (Fig. 1A). PEG treatment resulted in weight loss
in all 3 groups of PEG-treated mice, with the greatest change in weight
observed on the fifth day of PEG treatment and the mice regained most of
the weight five days after treatment (Fig. 1B). After either PEG,
clindamycin, or PEG and clindamycin treatment all mice were challenged
with 10\textsuperscript{3} \emph{C. difficile} 630 spores (Fig. 1A). All
treatments rendered mice susceptible to \emph{C. difficile}
colonization. However, PEG-treated mice remained colonized with \emph{C.
difficile} at high levels through thirty days post-infection (Fig. 1C).
In contrast, the clindamycin-treated mice cleared \emph{C. difficile}
within ten days post-infection. Surprisingly, PEG-treated mice were
still susceptible to \emph{C. difficile} infection after 10-days of
recovery from treatment although \emph{C. difficile} was not detectable
in most of the group in the initial five days post-infection (Fig. 1C).
From 9 days post-infection onward, the median \emph{C. difficile}
stabilized for the 10-day recovery group of PEG-treated mice and
remained high through 30 days post-infection (Fig. 1C). Thus, osmotic
laxative treatment alone was sufficient to render mice susceptible to
prolonged \emph{C. difficile} colonization and PEG-treated mice remained
susceptible for up to ten days post-treatment.

\textbf{5-day laxative treatment differentially disrupts the fecal
microbiota compared to clindamycin treatment.} Since osmotic laxatives
and clindamycin have previously been shown to disrupt the murine
microbiota (14--17), we hypothesized the different \emph{C. difficile}
colonization dynamics between mice treated with the osmotic laxative or
clindamycin were due to the two drugs having differential effects on the
microbiota. We profiled the stool microbiota over time by sequencing the
V4 region of the 16S rRNA gene to compare changes across treatment
groups. We found time and treatment group explained half of the observed
variation between fecal communities with most of the remaining variation
explained by interactions between treatment group and other experimental
variables including time, cage effects, and sequencing preparation
(PERMANOVA combined R\textsuperscript{2} = 0.95, \emph{P} \textless{}
0.001, Fig. 2A, Data Set S1, Sheet X). Cage effects refer to the
well-documented phenomenon that mice housed in the same cages have
similar microbial communities due to copraphagy (27), we tried to
minimize the impact of cage effects on our experiment by breaking up
cagemates when assigning mice to treatment groups and primarily housing
only two mice per cage. Importantly, although we conducted a total of 5
separate experiments, the experiment number and it's interaction with
treatment group was not one of the variables that significantly
explained the observed variation in fecal communities (Data Set S1,
Sheet X). Interestingly, none of the treatment groups recovered to their
baseline community structure either 10 or 30 days post-infection
suggesting other community features besides recovery to baseline were
responsible for the prolonged \emph{C. difficile} colonization in
PEG-treated mice (Fig. 2B).

@@

Next, we examined CDI severity by evaluating cecum and colon
histopathology (28) and found there was no difference in cecum and colon
scores between clindamycin and PEG-treated mice that were challenged
with \emph{C. difficile} at 4 days post-infection (Fig. 3D), the
timepoint typically examined in \emph{C. difficile} 630 challenged mice
(29). We also looked at 6 days post-infection because that was when we
started to see a large difference in C\emph{. difficile} colonization
levels between PEG- and clindamycin-treated mice (Fig. 1C). Although,
there was a slight difference in the colon between PEG and
clindamycin-treated mice, there was no difference in the cecum and the
overall score is still relatively low given that the max possible
summary score is 12 (Fig. 3E). Thus, although PEG treatment had a
profound impact on the mucosal microbiota, the impact of \emph{C.
difficile} 630 infection on the cecum and colon was similar between PEG
and clindamycin treated mice.

\textbf{\emph{C. difficile} challenge does not have a synergistic
disruptive effect on the microbiota of PEG-treated mice} Because
\emph{C. difficile} itself can have an impact on the microbiota (30), we
also sequenced the tissue and stools of mock-challenged clindamycin and
5-day PEG treated mice. Examining the stools of the mock-challenged mice
revealed similar bacterial disruptions as the \emph{C. difficile}
challenged mice (Fig. S2A-C). Similarly, there was no difference between
the tissues of mock and \emph{C. difficile} challenged mice (Fig.
S2D-F). Thus, most of the microbiota alterations we observed in the
PEG-treated mice were a result of the laxative and not an interaction
between the laxative and \emph{C. difficile}.

@@

\textbf{Five-day post-infection community data can predict mice that
will have prolonged \emph{C. difficile} colonization} After identifying
bacteria associated with the 5-day, 1-day and post-CDI 1-day PEG
treatments, we decided to examine the taxa that were influencing
prolonged \emph{C. difficile} colonization. We trained 3 types of
machine learning models (random forest, logistic regression, and support
vector machine) with input bacterial community data from 5 days
post-infection to predict whether the mice were colonized with \emph{C.
difficile} 10 days post-infection. We chose 5 days post-infection
because that was the earliest timepoint where we would see a treatment
effect in the mice that were given 1-day PEG treatment three days after
\emph{C. difficile} challenge and then administered an FMT or PBS
gavage. The random forest model had the highest performance (AUROC =
0.90, Data Set S1, Sheet X), so we next performed permutation importance
to examine the bacteria that were driving performance. We selected the
top 10 bacteria contributing to our models performance (Fig. 7A) and
examined their relative abundance at 5 days post-infection, the
timepoint used to predict C. difficile colonization status on day 10
(Fig. 7B). Next, we focused on the 5 genera that were \textgreater{} 1
\% relative abundance in either the cleared or colonized mice and
examined the relative abundance dynamics of these bacteria over time. We
found \emph{, }, and \_ tended to have a higher relative abundance and
\textbf{, } had a lower relative abundance in the mice with prolonged
colonization compared to the mice that cleared \emph{C. difficile} (Fig.
7C). Previous work examining the impact of PEG on the murine microbiota
found that PEG treatment resulted in the permanent loss of S24-7, also
known as \emph{Muribaculum intestinale} (16). We decided to check our
\emph{Porphyromonadaceae} OTUs because \emph{Muribaculum intestinale} is
known to be classified as \emph{Porphyromonadaceae} by the Ribosomal
Database Project (RDP) database (31), \emph{Porphyromonadaceae} was a
top feature in the random forest model predicting prolonged \emph{C.
difficile} colonization, and had a high relative abundance in the
communities of mice that cleared \emph{C. difficile} within 10 days. We
identified 4 OTUs that had at least (insert minimum percent identity) to
\emph{Muribaculum intestinale} and examined their abundance in mice that
either cleared or were still colonized with \emph{C. difficile} at 10
days post-infection (Fig. S4).

@@

  \begin{itemize}
  \tightlist
  \item
    What's known regarding laxatives and susceptibility to CDIs
  \item
    Clinical trial of PEG, results never posted (32)
  \item
    Relevance to human FMTs? Unclear what the best administration route
    is because there have been no studies designed to evaluate the best
    administration route for FMTs.
  \end{itemize}
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  Histopathology (33)
\end{itemize}

@@

\leavevmode\hypertarget{ref-Schubert2014}{}%
12. \textbf{Schubert AM}, \textbf{Rogers MAM}, \textbf{Ring C},
\textbf{Mogle J}, \textbf{Petrosino JP}, \textbf{Young VB},
\textbf{Aronoff DM}, \textbf{Schloss PD}. 2014. Microbiome data
distinguish patients with clostridium difficile infection and non-c.
Difficile-associated diarrhea from healthy controls. mBio \textbf{5}.
doi:\href{https://doi.org/10.1128/mbio.01021-14}{10.1128/mbio.01021-14}.

@@

\leavevmode\hypertarget{ref-Olson2013}{}%
18. \textbf{Olson A}, \textbf{Diebel LN}, \textbf{Liberati DM}. 2013.
Effect of host defenses on clostridium difficile toxininduced intestinal
barrier injury. Journal of Trauma and Acute Care Surgery
\textbf{74}:983--990.
doi:\href{https://doi.org/10.1097/ta.0b013e3182858477}{10.1097/ta.0b013e3182858477}.

\leavevmode\hypertarget{ref-Diebel2014}{}%
19. \textbf{Diebel LN}, \textbf{Liberati DM}. 2014. Reinforcement of the
intestinal mucus layer protects against clostridium difficile intestinal
injury in~vitro. Journal of the American College of Surgeons
\textbf{219}:460--468.
doi:\href{https://doi.org/10.1016/j.jamcollsurg.2014.05.005}{10.1016/j.jamcollsurg.2014.05.005}.

\leavevmode\hypertarget{ref-vanNood2013}{}%
20. \textbf{Nood E van}, \textbf{Vrieze A}, \textbf{Nieuwdorp M},
\textbf{Fuentes S}, \textbf{Zoetendal EG}, \textbf{Vos WM de},
\textbf{Visser CE}, \textbf{Kuijper EJ}, \textbf{Bartelsman JFWM},
\textbf{Tijssen JGP}, \textbf{Speelman P}, \textbf{Dijkgraaf MGW},
\textbf{Keller JJ}. 2013. Duodenal infusion of donor feces for
RecurrentClostridium difficile. New England Journal of Medicine
\textbf{368}:407--415.
doi:\href{https://doi.org/10.1056/nejmoa1205037}{10.1056/nejmoa1205037}.

\leavevmode\hypertarget{ref-Razik2017}{}%
21. \textbf{Razik R}, \textbf{Osman M}, \textbf{Lieberman A},
\textbf{Allegretti JR}, \textbf{Kassam Z}. 2017. Faecal microbiota
transplantation for clostridium difficile infection: A multicentre study
of non-responders. Medical Journal of Australia \textbf{207}:159--160.
doi:\href{https://doi.org/10.5694/mja16.01452}{10.5694/mja16.01452}.

\leavevmode\hypertarget{ref-Postigo2012}{}%
22. \textbf{Postigo R}, \textbf{Kim JH}. 2012. Colonoscopic versus
nasogastric fecal transplantation for the treatment of clostridium
difficile infection: A review and pooled analysis. Infection
\textbf{40}:643--648.
doi:\href{https://doi.org/10.1007/s15010-012-0307-9}{10.1007/s15010-012-0307-9}.

\leavevmode\hypertarget{ref-Liacouras1996}{}%
23. \textbf{Liacouras CA}, \textbf{Piccoli DA}. 1996. Whole-bowel
irrigation as an adjunct to the treatment of chronic, relapsing
clostridium difficile colitis. Journal of Clinical Gastroenterology
\textbf{22}:186--189.
doi:\href{https://doi.org/10.1097/00004836-199604000-00007}{10.1097/00004836-199604000-00007}.

\leavevmode\hypertarget{ref-Schubert2015}{}%
24. \textbf{Schubert AM}, \textbf{Sinani H}, \textbf{Schloss PD}. 2015.
Antibiotic-induced alterations of the murine gut microbiota and
subsequent effects on colonization resistance against \emph{Clostridium
difficile}. mBio \textbf{6}.
doi:\href{https://doi.org/10.1128/mbio.00974-15}{10.1128/mbio.00974-15}.

\leavevmode\hypertarget{ref-Jenior2017}{}%
25. \textbf{Jenior ML}, \textbf{Leslie JL}, \textbf{Young VB},
\textbf{Schloss PD}. 2017. \emph{Clostridium difficile} colonizes
alternative nutrient niches during infection across distinct murine gut
microbiomes. mSystems \textbf{2}.
doi:\href{https://doi.org/10.1128/msystems.00063-17}{10.1128/msystems.00063-17}.

\leavevmode\hypertarget{ref-Tomkovich2020}{}%
26. \textbf{Tomkovich S}, \textbf{Stough JMA}, \textbf{Bishop L},
\textbf{Schloss PD}. 2020. The initial gut microbiota and response to
antibiotic perturbation influence clostridioides difficile clearance in
mice. mSphere \textbf{5}.
doi:\href{https://doi.org/10.1128/msphere.00869-20}{10.1128/msphere.00869-20}.

\leavevmode\hypertarget{ref-Nguyen2015}{}%
27. \textbf{Nguyen TLA}, \textbf{Vieira-Silva S}, \textbf{Liston A},
\textbf{Raes J}. 2015. How informative is the mouse for human gut
microbiota research? Disease Models \& Mechanisms \textbf{8}:1--16.
doi:\href{https://doi.org/10.1242/dmm.017400}{10.1242/dmm.017400}.

\leavevmode\hypertarget{ref-Reeves2011}{}%
28. \textbf{Reeves AE}, \textbf{Theriot CM}, \textbf{Bergin IL},
\textbf{Huffnagle GB}, \textbf{Schloss PD}, \textbf{Young VB}. 2011. The
interplay between microbiome dynamics and pathogen dynamics in a murine
model of \emph{Clostridium difficile} infection \textbf{2}:145--158.
doi:\href{https://doi.org/10.4161/gmic.2.3.16333}{10.4161/gmic.2.3.16333}.

\leavevmode\hypertarget{ref-Dieterle2020}{}%
29. \textbf{Dieterle MG}, \textbf{Putler R}, \textbf{Perry DA},
\textbf{Menon A}, \textbf{Abernathy-Close L}, \textbf{Perlman NS},
\textbf{Penkevich A}, \textbf{Standke A}, \textbf{Keidan M},
\textbf{Vendrov KC}, \textbf{Bergin IL}, \textbf{Young VB}, \textbf{Rao
K}. 2020. Systemic inflammatory mediators are effective biomarkers for
predicting adverse outcomes in clostridioides difficile infection. mBio
\textbf{11}.
doi:\href{https://doi.org/10.1128/mbio.00180-20}{10.1128/mbio.00180-20}.

\leavevmode\hypertarget{ref-Jenior2018}{}%
30. \textbf{Jenior ML}, \textbf{Leslie JL}, \textbf{Young VB},
\textbf{Schloss PD}. 2018. \emph{Clostridium difficile} alters the
structure and metabolism of distinct cecal microbiomes during initial
infection to promote sustained colonization. mSphere \textbf{3}.
doi:\href{https://doi.org/10.1128/msphere.00261-18}{10.1128/msphere.00261-18}.

\leavevmode\hypertarget{ref-Vornhagen2020}{}%
31. \textbf{Vornhagen J}, \textbf{Bassis CM}, \textbf{Ramakrishnan S},
\textbf{Hein R}, \textbf{Mason S}, \textbf{Bergman Y}, \textbf{Sunshine
N}, \textbf{Fan Y}, \textbf{Timp W}, \textbf{Schatz MC}, \textbf{Young
VB}, \textbf{Simner PJ}, \textbf{Bachman MA}. 2020. A plasmid locus
associated with klebsiella clinical infections encodes a
microbiome-dependent gut fitness factor.
doi:\href{https://doi.org/10.1101/2020.02.26.963322}{10.1101/2020.02.26.963322}.

\leavevmode\hypertarget{ref-Dieterle2018}{}%
32. \textbf{Dieterle MG}, \textbf{Rao K}, \textbf{Young VB}. 2018. Novel
therapies and preventative strategies for primary and
recurrentClostridium difficileinfections. Annals of the New York Academy
of Sciences \textbf{1435}:110--138.
doi:\href{https://doi.org/10.1111/nyas.13958}{10.1111/nyas.13958}.

\leavevmode\hypertarget{ref-Theriot2011}{}%
33. \textbf{Theriot CM}, \textbf{Koumpouras CC}, \textbf{Carlson PE},
\textbf{Bergin II}, \textbf{Aronoff DM}, \textbf{Young VB}. 2011.
Cefoperazone-treated mice as an experimental platform to assess
differential virulence ofClostridium difficilestrains. Gut Microbes
\textbf{2}:326--334.
doi:\href{https://doi.org/10.4161/gmic.19142}{10.4161/gmic.19142}.
